---
title: "Lab 1"
author: "Nicole Schmidt"
date: "1/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#attach packages 

library(tidyverse)
library(here)
library(dplyr)
library(tmap)
library(sf)

sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))

```

### Basic wrangling reminders

Refresh some skills for data wrangling & summary statistics using functions in the `dplyr` package 

Find the top 5 highest observations of trees by legal_status. Do some wrangling, make a graph. 
```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% #count function is recognizing groups, finding counts, and puts it in a nice table (takes group by, summarise, n and uses all of it). Can have multiple levels of grouping as well. 
drop_na(legal_status) %>% #removes missing values; careful when doing this bc can be problematic.
rename(tree_count = n) %>%  # new name replaces old name
relocate(tree_count) %>% 
#allows you to move things around by class (character or class) or location, or columns) 
  slice_max(tree_count, n = 5)
```

Make a graph of those top 5 observations by legal status. 
```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count))+
  geom_col()+
  labs(x = "Legal Status", y = "Tree Count")+
  coord_flip()+
  theme_minimal()

```

### A few more data wrangling refresher examples 

Only want to keep observations (rows) for Blackwood Acacia Trees

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% #filter to keep observations if within the species variable the string is detected where that stric is x. not looking for a complete exact match but only looking for an observation where that string shows up. 
select(legal_status, date, latitude, longitude)

ggplot(data = blackwood_acacia, aes( x = longitude, y = latitude))+
  geom_point()
```

### tidyr::seperate() and unite() functions

Useful for combining or seperating columns 

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

Example: tidyr::unite()

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")
```

###Make some actual Maps of blackwood acacia trees in SF 

`st_as_as()` to convert latitude and longitude to spatial coordinates.

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c("longitude", "latitude"))

st_crs(blackwood_acacia_sp) = 4326 #sets coordinate reference system 

#graph using geom_sf
ggplot(data = blackwood_acacia_sp) +
  geom_sf(color = "darkgreen")

```


###Read in SF roads shapefile

```{r}
sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp")) # read in spatial shapfile 

st_transform(sf_map, 4326) # data will have same coordinate reference system 

ggplot(data = sf_map)+
  geom_sf()
```


Combine blackwood acacia tree observations and SF roads map

```{r}
ggplot()+
  geom_sf(data = sf_map, size = 0.1, color = "darkgray")+
  geom_sf(data = blackwood_acacia_sp, color = "red", size = 0.5)+
  theme_void()
```

now interactive map!!!
```{r}
tmap_mode("view")

tm_shape(blackwood_acacia_sp)+
  tm_dots()
```

